### 2023-08-07 pilot-publication/lab-book.Rmd
Separate functions to generate results

# Lab book for the TERN pilot publication

This file includes code and scripts used to experiment and explore with the TERN pilot data.

```{r setup}
library(pbapply)
library(openssl)
library(dotenv)
library(DataExplorer)
```

## Generate synthetic data

Because developing on the server is generally slow, a first step to speed up development was to generate a synthetic dataset that could be safely used locally. Check [main.R](./main.R) to make sure that `data` is in the correct state before creating the synthetic data using the code below.

```{r, synthetic-data-generation}
## Generate synthetic data. For each column, sample from the same distribution as the original data, but do not keep correlations present in the orginal data.
synth.data <- lapply(data, function(column) {
    if (is.factor(column)) {
        return (as.factor(sample(levels(column), size = length(column), replace = TRUE)))
    } else if (is.character(column)) {
        return (as.character(sample(column, size = length(column), replace = TRUE)))
    } else if (is.numeric(column)) {
        return (rnorm(n = length(column), mean = mean(column, na.rm = TRUE), sd = sd(column, na.rm = TRUE)))
    } else if (is.logical(column)) {
        return (sample(c(TRUE, FALSE), size = length(column), replace = TRUE))
    } else if (is.POSIXct(column) | is.POSIXt(column)) {
        return (as.POSIXct(sample(as.numeric(column), size = length(column), replace = TRUE), origin = "1970-01-01"))
    } else {
        return (rep(NA, length(column)))
    }
}) %>% as.data.frame()

## Replace values in columns with character values with random strings
synth.data[, sapply(synth.data, is.character)] <- lapply(synth.data[, sapply(synth.data, is.character)], function(column) {
    n.unique <- length(unique(column))
    new.values <- replicate(n.unique, paste(sample(letters, size = 10, replace = TRUE), collapse = ""))
    new.values <- rep(new.values, length.out = length(column))
    return (new.values)
})

## Replace values in any variable with id in the name random numbers
id.vars <- c("X_id", "id__reg_hospital_id", "id__assigned_patient_id", "meta__instanceID", "meta__deprecatedID", "formhub__uuid", "X_xform_id_string", "X_uuid")
synth.data[, id.vars] <- lapply(synth.data[, id.vars], function(column) {
    n.unique <- length(unique(column))
    new.values <- rep(sample(1:1e6, size = n.unique, replace = FALSE), length.out = length(column))
    return (new.values)
})

## Check that no rows in the original data are present in the synthetic data
stopifnot(!any(pbapply(data, 1, function(row) {
    any(apply(synth.data, 1, function(synth.row) {
        all(row == synth.row, na.rm = TRUE)
    }))
})))

## Put labels back
synth.data <- synth.data %>% labelled::copy_labels_from(data)

## Create data report
DataExplorer::create_report(synth.data, report_title = "Synthetic data report", output_file = "synth-data-report.html")

## Encrypt synthetic data before writing to file
key <- sha256(charToRaw(Sys.getenv("SYNTH_DATA_KEY")))
encrypted.synth.data <- aes_cbc_encrypt(serialize(synth.data, NULL), key = key)

## Write synthetic data to file
saveRDS(encrypted.synth.data, file = "data/synth-data.rds")

## Encrypt and write codebook to file, but first remove the settings entry
codebook$settings <- NULL
encrypted.codebook <- aes_cbc_encrypt(serialize(codebook, NULL), key = key)
saveRDS(encrypted.codebook, file = "data/codebook.rds")
```

To develop locally both the data and the codebook need to be loaded.

```{r load-data}
key <- sha256(charToRaw(Sys.getenv("SYNTH_DATA_KEY")))
data <- unserialize(aes_cbc_decrypt(readRDS("data/synth-data.rds"), key = key))
codebook <- unserialize(aes_cbc_decrypt(readRDS("data/codebook.rds"), key = key))
```

Probably a good idea to wrap the generation of these results into a function, and introduce a test argument, to prevent us from introducing local results in the main manuscript.

```{r basic-results}
get_basic_results <- function(test = FALSE) {
    results <- list()
    ## Define basic results
    arrival.dates <- data %>% pull(incident__date_of_arrival) %>% as.Date()
    results$start.date <- arrival.dates %>% min() %>% format_date()
    results$end.date <- arrival.dates %>% max() %>% format_date()
    n.no.consent <-  list("11542" = 40,
                          "44805" = 10,
                          "55356" = 43,
                          "78344" = 3,
                          "95846" = 9, 
                          "88456" = 0, # To be updated
                          "10263" = 2)
    results$icc <- estimate_icc("outcomes__discharge_alive", "id__reg_hospital_id", data)
    results$n.patients <- nrow(data)
    results$n.atls.residents <- 4 + 2 # The total number of residents trained in ATLS, per ATLS centre
    results$n.ptc.residents <- 9 + 6 # The total number of residents trained in PTC, per centre
    results$n.residents <- with(results, n.atls.residents + n.ptc.residents)
    results$n.centres <-  data %>% pull(id__reg_hospital_id) %>% unique() %>% length()
    results$n.atls <- sum(data$arm == "ATLS")
    results$n.ptc <- sum(data$arm == "PTC")
    results$n.control <- sum(data$arm == "Standard care")
    results$n.females <- with(data, sum(patinfo__pt_gender == "Female"))
    results$p.females <- round(results$n.females/nrow(data) * 100)
    results$median.age <- median(data$patinfo__pt_age, na.rm = TRUE)
    results$iqr.age <- get_iqr(data$patinfo__pt_age)
    results$median.iss <- median(data$riss, na.rm = TRUE)
    results$iqr.iss <- get_iqr(data$riss)
    results$median.niss <- median(data$niss, na.rm = TRUE)
    results$iqr.niss <- get_iqr(data$niss)
    results$n.admitted <- with(data, sum(interventions__admitted == "Yes"))
    results$p.admitted <- round(results$n.admitted/nrow(data) * 100)
    return (results)
}
```

Cutting code from the main script and putting it here for future reference:

```{r}
## Calculate outcome results
n.m30d <- with(data, sum(outcomes__alive_after_30_days == "Yes", na.rm = TRUE))
p.m30d <- round(n.m30d/nrow(data) * 100)
arms.n.m30d.list <- lapply(setNames(arms.data.list, nm = paste0("m30d.", names(arms.data.list))),
                           function(arms.data)
                               with(arms.data,
                                    sum(outcomes__alive_after_30_days == "Yes", na.rm = TRUE)))
arms.p.m30d.list <- mapply(arms.n.m30d.list, arms.data.list,
                           FUN = function(count, arms.data) list(round(count/nrow(arms.data) * 100)))
names(arms.n.m30d.list) <- paste0("n.", names(arms.n.m30d.list))
names(arms.p.m30d.list) <- paste0("p.", names(arms.p.m30d.list))
attach(arms.n.m30d.list)
attach(arms.p.m30d.list)
arr.atls.ptc <- p.m30d.atls - p.m30d.ptc
arr.atls.control <- p.m30d.atls - p.m30d.control
arr.ptc.control <- p.m30d.ptc - p.m30d.control
rr.atls.ptc <- round(p.m30d.atls/p.m30d.ptc, 2)
rr.atls.control <- round(p.m30d.atls/p.m30d.control, 2)
rr.ptc.control <- round(p.m30d.ptc/p.m30d.control, 2)
p.missing.in.hospital.mortality <- round(sum(is.na(data$outcomes__discharge_alive))/n.patients * 100)

## Estimate composite outcome
in.hospital.mortality <- data$outcomes__discharge_alive == "Yes"
data$in.hospital.mortality <- in.hospital.mortality
icc.in.hospital.mortality <- estimate_icc("in.hospital.mortality", "id__reg_hospital_id", data)
labelled::var_label(data$in.hospital.mortality) <- "In-hospital mortality"
confined.to.bed <- data$outcomes__eq5dm == "I am confined to bed"
data$confined.to.bed <- confined.to.bed
labelled::var_label(data$confined.to.bed) <- "Confined to bed"
extreme.pain.discomfort <- data$outcomes__eq5dpd == "I have extreme pain or discomfort"
data$extreme.pain.discomfort <- extreme.pain.discomfort
labelled::var_label(data$extreme.pain.discomfort) <- "Extreme pain or discomfort"
unable.bath.dress <- data$outcomes__eq5dsc == "I am unable to bathe or dress myself"
data$unable.bath.dress <- unable.bath.dress
labelled::var_label(data$unable.bath.dress) <- "Unable to bathe or dress oneself"
unable.usual.activities <- data$outcomes__eq5dua == "I am unable to perform my usual activities"
data$unable.usual.activities <- unable.usual.activities
labelled::var_label(data$unable.usual.activities) <- "Unable to perform usual activities"
composite.outcome <- in.hospital.mortality |
    confined.to.bed |
    extreme.pain.discomfort |
    unable.bath.dress |
    unable.usual.activities
data$composite.outcome <- composite.outcome
labelled::var_label(data$composite.outcome) <- "Composite endpoint"
n.composite.outcome <- sum(composite.outcome, na.rm = TRUE)
p.composite.outcome <- round(n.composite.outcome/nrow(data) * 100)
icc.composite.outcome <- estimate_icc("composite.outcome", "id__reg_hospital_id", data)

## Resident outcomes
median.confidence <- median(as.numeric(data$resident__res_comfort), na.rm = TRUE)
iqr.confidence <- quantile(as.numeric(data$resident__res_comfort), probs = c(0.25, 0.75), na.rm = TRUE) %>%
    paste0(collapse = "-")

## Feasibility outcomes

## List of potentially eligible patient participants per site, replace
## IDs with actual IDs once known and replace figures with actual
## figures once known
n.potentially.eligible = list(n.potentially.eligible.centre.1 = 999,
                              n.potentially.eligible.centre.2 = 999,
                              n.potentially.eligible.centre.3 = 999,
                              n.potentially.eligible.centre.4 = 999,
                              n.potentially.eligible.centre.5 = 999,
                              n.potentially.eligible.centre.6 = 999,
                              n.potentially.eligible.centre.7 = 999)
recruitment.rate.patients <- round(n.patients/do.call(sum, n.potentially.eligible) * 100)
n.eligible.residents <- 21
recruitment.rate.residents <- round(n.residents/n.eligible.residents * 100)
n.lost.to.follow.up <- sum(is.na(data$outcomes__alive_after_30_days))
rate.lost.to.follow.up <- round(n.lost.to.follow.up/n.patients * 100)
n.atls.residents.passed.first.attempt <- 2 + 2 # The number of residents who had passed ATLS after the first attempt, per centre
n.atls.residents.passed.second.attempt <- 4 + 2 # The number of residents who had passed ATLS after the second attempt, per centre
rate.pass <- round((n.atls.residents.passed.second.attempt + n.ptc.residents)/n.residents * 100)
rates.missing.data <- unlist(lapply(data, function(column) round(sum(is.na(column)/length(column) * 100))))
min.missing.data <- min(rates.missing.data)
max.missing.data <- max(rates.missing.data)

## Calculate varying cluster sizes
cluster.sizes <- sapply(centre.ids, function(centre.id) nrow(data[data$id__reg_hospital_id == centre.id, ]))
sd.cluster.size <- sd(cluster.sizes)
mean.cluster.size <- mean(cluster.sizes)
ratio.sd.mean.cluster.size <- sd.cluster.size/mean.cluster.size
```