### 2024-10-14 pilot-publication/supplementary-materials.Rmd
Compila manuscript and supplementary materials

---
title: Supplementary Materials
subtitle: "Effect of Trauma Life Support Training on Patient Outcomes: A Pilot Cluster Randomised Trial"
output:
    bookdown::word_document2:
      number_sections: false
      reference_docx: "reference.docx"
---

```{r source, echo = FALSE, include = FALSE}
source("main.R")
attach(results)
```

```{r, child = "outcome-tables.Rmd"}
```