### 2024-10-18 pilot-publication/supplementary-outcome-tables.Rmd

\pagebreak

# Results with 95% confidence level

## Table S2 Outcomes in all patients during the entire study period, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Table S3 Outcomes in all patients before training, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Table S4 Outcomes in all patients after training, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Table S5 Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Table S6 Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Table S7 Absolute and relative differences in outcomes after training, comparing standard care with ATLS

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Table S8 Absolute and relative differences in outcomes after training, comparing standard care with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Table S9 Absolute and relative differences in outcomes after training, comparing ATLS with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Table S10 Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Table S11 Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Table S12 Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak