### 2024-10-19 pilot-publication/supplementary-outcome-tables.Rmd
Compile new version of manuscript

\pagebreak

# Results with 95% confidence level

## Table S2. Outcomes in all patients during the entire study period, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Table S3. Outcomes in all patients before training, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Table S4. Outcomes in all patients after training, by treatment arm

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Table S5. Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Table S6. Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Table S7. Absolute and relative differences in outcomes after training, comparing standard care with ATLS

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Table S8. Absolute and relative differences in outcomes after training, comparing standard care with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Table S9. Absolute and relative differences in outcomes after training, comparing ATLS with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Table S10. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Table S11. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Table S12. Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak