### 2024-11-20 outcome-tables.Rmd
Move everything to top directory

\pagebreak

# Results with 75% confidence intervals (CI)

## Outcomes in all patients during the entire study period, by treatment arm(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients before training, by treatment arm(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients after training, by treatment arm(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with ATLS(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with PTC(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing ATLS with PTC(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC(75% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak

\pagebreak

# Results with 85% confidence intervals (CI)

## Outcomes in all patients during the entire study period, by treatment arm(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients before training, by treatment arm(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients after training, by treatment arm(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with ATLS(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with PTC(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing ATLS with PTC(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC(85% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak

\pagebreak

# Results with 95% confidence intervals (CI)

## Outcomes in all patients during the entire study period, by treatment arm(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients before training, by treatment arm(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Outcomes in all patients after training, by treatment arm(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with ATLS(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing standard care with PTC(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in outcomes after training, comparing ATLS with PTC(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC(95% CI)

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak