### 2024-11-20 table-test.tex

\documentclass{article}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{graphicx}

\begin{document}

\begingroup
\fontsize{12.0pt}{14.4pt}\selectfont
\setlength{\LTpost}{0mm}
\begin{longtable}{lcccccc}
\toprule
 &  &  &  & \multicolumn{3}{c}{\textbf{Differences}} \\ 
\cmidrule(lr){5-7}
\textbf{Outcome} & \textbf{Standard care N = 161\textsuperscript{\textit{1}}} & \textbf{ATLS N = 28\textsuperscript{\textit{1}}} & \textbf{PTC N = 73\textsuperscript{\textit{1}}} & \textbf{Standard care vs. ATLS} & \textbf{Standard care vs. PTC} & \textbf{ATLS vs. PTC} \\ 
\midrule\addlinespace[2.5pt]
30 day mortality & 22 (16\%) & 1 (3.8\%) & 3 (4.9\%) & 12\% & 11\% & -1.1\% \\ 
Unknown & 26 & 2 & 12 &  &  &  \\ 
In-hospital mortality & 19 (12\%) & 1 (3.7\%) & 3 (4.1\%) & 8.2\% & 7.8\% & -0.41\% \\ 
Unknown & 2 & 1 & 0 &  &  &  \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}n (\%)\\
\end{minipage}
\endgroup

\end{document}