### 2025-01-08 cover-letter.qmd
Restore directory state to BMJ Open version

---
title: "Cover letter"
format: pdf
---

\vspace{-2cm}

Dear Editor,

I am submitting our manuscript, “Feasibility of a Cluster Randomised Trial on the Effect of Trauma Life Support Training: A Pilot Study” for consideration for publication in JAMA Surgery.

This study evaluates the feasibility of conducting a full-scale trial to assess the impact of trauma life support training programs on patient outcomes. The findings include consent rates, data quality, and effect size estimates, which are essential for planning future research. We believe this work addresses an important gap in evidence for trauma care and will be of interest to your readers.

This manuscript has not been published or submitted to any journal, but it has been posted on the pre-print server medRxiv: https://doi.org/10.1101/2024.03.13.24304236. No other related papers from this study have been published, posted, or submitted.

I am the corresponding author and can be contacted at:

---

\begin{center}
Martin Gerdin Wärnberg, MD, PhD

Department of Global Public Health

Karolinska Institutet

171 77 Stockholm, Sweden

Email: martin.gerdin@ki.se

Phone: +46 708 53 95 98
\end{center}

---

Thank you for considering our manuscript.

Sincerely,

Martin Gerdin Wärnberg, MD, PhD