### 2025-06-27 supplementary-materials.Rmd
Continue to work on response to reviewers

---
## Rscript -e "rmarkdown::render('supplementary-materials.Rmd')"
## libreoffice --headless --convert-to pdf supplementary-materials.docx
title: Supplementary Materials
subtitle:"Feasibility of a Cluster Randomised Trial on the Effect of Trauma Life Support Training: A Pilot Study in India"
output:
    bookdown::word_document2:
      number_sections: false
      reference_docx: "reference.docx"
      toc: true
---

```{r source, echo = FALSE, include = FALSE}
source("main.R")
attach(results)
```

# Protocol Deviations

## Trial Registration

We intended to register our trial with Clinical Trials Registry - India (CTRI), but because of a communication error in the management group this was only attempted after the trial has started including patients, and CTRI only accept prospective registrations. We intend to register the full scale trial with both ClinicalTrials.gov and CTRI.

## Reporting

In the original protocol, we also aimed to estimate probable effect sizes and other measures needed for the sample size calculations of a full-scale trial, but we revised this aim in the light of current guidance on the conduct and reporting of pilot trials

## Number of Participating Centres

We recruited seven centres instead of six and therefore assigned two centres each to the intervention arms and three centres to the control arm.

## Resident Participants

We included emergency medicine residents in addition to surgical residents.

## Periodic surveys of residents

We did not distribute periodic surveys to the participating residents but discussed challenges and suggestions that they had regarding the scheduling or implementation of the training programs.

## Follow up of residents

We stated that resident participants would be followed up 30 days after training, but revised this to follow them up after the end of the study period.

## Data collection from records

We decided to extract data from medical records only for a convenience sample of patients to reduce the research officers' workload.

## Selection of units for training

We planned to use simple random sampling to select units if there were more than two eligible units in a hospital but instead the hospital principal investigator decided which units to train.

## Timing of resident consent

We had initially planned to ask residents for consent before randomisation, but because of logistical issues the units were only finalised after the hospitals had been randomised. Residents were therefore approached for consent after randomisation but before training.

## Analysis level of feasibility outcomes

We had planned to analyse feasibility outcomes on both an overall and individual cluster level, but we only analysed them on an overall level, because the sample sizes in individual clusters were too small to generate meaningful results.

## Patient outcomes across subgroups

Because of small numbers in the pre-specified subgroups men, women, blunt multisystem trauma, penetrating trauma, shock, severe traumatic brain injury and elderly we decided to report only descriptive data on these subgroups.

# Table S1. Cluster characteristics

| Cluster                                                       | City       | Beds | Type    | Study period | 
+---------------------------------------------------------------+------------+------+---------+--------------+
| Dr R N Cooper Municipal General Hospital                      | Mumbai     |      | Public  |              |
| Seth GS Medical College and KEM Hospital                      | Mumbai     |      | Public  |              |
| Lokmanya Tilak Municipal Medical College and General Hospital | Mumbai     |      | Public  |              |
| MEDICA Superspeciality Hospital                               | Kolkata    |      | Private |              |
| Medical College                                               | Kolkata    |      | Public  |              |
| Nilratan Sircar Medical College & Hospital                    | Kolkata    |      | Public  |              |
| Postgraduate Institute of Medical Education and Research      | Chandigarh |      | Public  |              |
+---------------------------------------------------------------+------------+------+---------+--------------+

# Statistical methods for additional analyses according to the original protocol

We used an empty generalised linear mixed model <!-- see [@Sommet2017] --> to estimate the intracluster correlation coefficient. We compared patients outcomes in all possible combinations of trial arms. In each combination we compared both differences in final values and differences in change from baseline. For the intervention arms the change from baseline was calculated as the difference between the one month period of data collection before the training was undertaken and the three month period after the training. For the control arm the data collection period was four months and the difference from baseline was calculated as the difference between the first one month and the following three months.

Within each combination of trial arms we had planned to conduct subgroup analyses of men, women, blunt multisystem trauma, penetrating trauma, shock (systolic blood pressure ≤ 90 mmHg), severe traumatic brain injury, and elderly (≥65 years) [@hornor_quality_2018]. These subgroups were however too small to allow for meaningful analyses, and are therefore reported descriptively. We calculated both absolute and relative differences for each comparison, along with 75, 85, and 95% confidence intervals. We used an empirical bootstrap procedure with 1000 draws to estimate these confidence intervals. In the interest of space, only the 95% confidence intervals are presented for all comparisons in Supplementary Tables. The remaining results are available from the corresponding author on request.

# Estimated intracluster correlation coefficient

The intracluster correlation coefficients was `r icc.m30d` for 30-day mortality and `r icc.inhosp` for in-hospital mortality.

# Table S2. Extended sample characteristics

```{r overall-table, echo = FALSE}
combined.sample.characteristics.table
```

```{r, child = "supplementary-outcome-tables.Rmd"}
```