### 2025-06-27 supplementary-outcome-tables.Rmd

\pagebreak

# Results with 95% confidence intervals (CI)

## Table S3. Outcomes in all patients during the entire study period, by treatment arm (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients during the entire study period, by treatment arm"]]
```

\pagebreak

## Table S4. Outcomes in all patients before training, by treatment arm (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients before training, by treatment arm"]]
```

\pagebreak

## Table S5. Outcomes in all patients after training, by treatment arm (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients after training, by treatment arm"]]
```

\pagebreak

## Table S6. Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arm (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]
```

\pagebreak

## Table S7. Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arm (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]
```

\pagebreak

## Table S8. Absolute and relative differences in outcomes after training, comparing standard care with ATLS (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]
```

\pagebreak

## Table S9. Absolute and relative differences in outcomes after training, comparing standard care with PTC (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]
```

\pagebreak

## Table S10. Absolute and relative differences in outcomes after training, comparing ATLS with PTC (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]
```

\pagebreak

## Table S11. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]
```

\pagebreak

## Table S12. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]
```

\pagebreak

## Table S13. Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC (95% CI)

```{r, echo = FALSE} 
outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]
```

\pagebreak