### 2025-07-05 manuscript.Rmd

---
## compile with:
## Rscript -e "rmarkdown::render('manuscript.Rmd', output_format = 'all')"
## libreoffice --headless --convert-to pdf manuscript.docx
title: "Feasibility of a Cluster Randomised Trial on the Effect of Trauma Life Support Training: A Pilot Study in India"
author: Trauma life support training Effectiveness Research Network (TERN) collaborators
bibliography: bibliography.bib
csl: nature.csl
output:
    bookdown::word_document2:
      toc: false
      number_sections: false
      reference_docx: "reference.docx"
#    md_document: default
#    bookdown::pdf_document2:
#      toc:false
#      number_sections:false
---

<!-- Source main code to make global objects available in the text -->
```{r source, echo = FALSE, include = FALSE}
source("main.R")
attach(results)
```

## Trial registration
This pilot study was registered with ClinicalTrials.gov (reg. no NCT05417243). We intended to also register this pilot trial with Clinical Trials Registry - India, but this was not done because of a communication error in the management group.

## Corresponding author

##### Martin Gerdin Wärnberg, MD, PhD (MGW) `r insert_orcid_link("0000-0001-6069-4794")`

Department of Global Public Health  
Karolinska Institutet  
171 77 Stockholm  
Sweden

martin.gerdin@ki.se  
+46 708 53 95 98 (mobile)figure

## Authors (in alphabetical order by last name)

##### Debojit Basak, MS `r insert_orcid_link("0000-0002-8378-9689")` <!-- Accepted, debojit88@gmail.com -->
Project Research Scientist, Institute of Post Graduate Medical Education and Research, Kolkata, India.

##### Johanna Berg, MD (JB) `r insert_orcid_link("0000-0001-7553-7337")` <!-- Accepted, johanna.berg@ki.se -->
Department of Global Public Health, Karolinska Institutet, Stockholm, Sweden.

Emergency Medicine, Department of Internal and Emergency Medicine, Skåne University Hospital, Malmö, Sweden.

##### Shamita Chatterjee, MBBS, MS (SC) `r insert_orcid_link("0000-0002-9460-108X")` <!-- Accepted, dr_shamita@yahoo.com -->
Department of Surgery, Institute of Post Graduate Medical Education and Research, Kolkata, India.

##### Li Felländer-Tsai, MD, PhD (LFT) `r insert_orcid_link("0000-0003-0693-6080")` <!-- Accepted -->
Division of Orthopaedics and Biotechnology, Department of Clinical Science Intervention and Technology (CLINTEC), Karolinska Institutet, Stockholm, Sweden.

Department of Reconstructive Orthopedics, Karolinska University Hospital, Stockholm, Sweden.

##### Geeta Ghag (GG), MBBS, MS `r insert_orcid_link("0000-0001-5561-362X")` <!-- Accepted, geetaghag@hotmail.com -->
Department of Surgery, HBT Medical College And Dr. R N Cooper Municipal General Hospital, Mumbai, India.

##### Catherine Juillard, MD, MPH (CJ) `r insert_orcid_link("0000-0003-0847-8933")` <!-- Accepted, cjuillard@mednet.ucla.edu -->
Division of General Surgery, Department of Surgery, David Geffen School of Medicine at UCLA, Los Angeles, California, United States of America.

##### Monty Khajanchi, MBBS, MS (MK) `r insert_orcid_link("0000-0002-0898-6391")` <!-- Accepted, monta32@gmail.com -->
Seth G. S. Medical College and K.E.M. Hospital, Mumbai, India.

##### Tamal Khan, MSc (TK) <!-- Accepted, tamal.khan440@gmail.com -->
Senior Research fellow, All India Institute of Medical Sciences, New Delhi, India.

##### Vipul Nandu, MBBS, MS (VN) `r insert_orcid_link("0000-0002-3366-5610")`  <!-- Accepted, dr.vipulnandu@gmail.com -->
Department of Surgery, HBT Medical College And Dr. R N Cooper Municipal General Hospital, Mumbai, India.

##### Nobhojit Roy, MBBS, MS, PhD (NR) `r insert_orcid_link("0000-0003-2022-7416")` <!-- Accepted, nroy@georgeinstitute.org.in -->
Program for Global Surgery and Trauma, The George Institute of Global Health, New Delhi, India.

Department of Global Public Health, Karolinska Institutet, Stockholm, Sweden.

##### Rajdeep Singh, MBBS, MS (RS) `r insert_orcid_link("0000-0001-6593-2624")` <!-- Accepted, rajdeep.mamc@gmail.com -->
Department of Surgery, Maulana Azad Medical College, New Delhi, India.

##### Kapil Dev Soni, MBBS, MD (KDS) `r insert_orcid_link("0000-0003-1214-4119")` <!-- Accepted, kdsoni111@gmail.com -->
Critical and Intensive Care, JPN Apex Trauma Center, All India Institute of Medical Sciences, New Delhi, India.

##### Lovisa Strömmer, MD, PhD (LS) `r insert_orcid_link("0000-0001-5424-7111")` <!-- Accepted, lisa.strommer@ki.se -->
Division of Surgery and Oncology, Department of Clinical Science, Intervention and Technology (CLINTEC), and Department of Global Public Health, Karolinska Institutet, Stockholm, Sweden.

## Non-author contributors

<!--
##### Prashant Bhandarkar (PB)
WHO collaboration Centre for Research in Surgical Care Delivery in LMIC, Mumbai, India.
-->

##### Anirban Chatterjee (AC) <!-- Accepted -->
Department of Orthopaedic Sciences, Medica Superspecialty Hospital, Kolkata, India.

<!--
##### Chintamani (C)
Department of Surgery, Vardhman Mahavir Medical College and Safdarjung Hospital, New Delhi, India.
-->

<!--
##### Anita Gadgil (AG)
WHO collaboration Centre for Research in Surgical Care Delivery in LMIC, Mumbai, India.
-->

##### Marie Hasselberg (MH) <!-- Accepted -->
Department of Global Public Health, Karolinska Institutet, Stockholm, Sweden.

<!--
##### Deepa Kizhakke Veetil (DKV)
Department of Surgery, Manipal Hospital, Dwarka, New Delhi.
-->

<!--
##### Vineet Kumar (VK)
Department of Surgery, Lokmanya Tilak Municipal Medical College and General Hospital, Mumbai, India.
-->

##### Debabrata Kundu (DK) <!-- Accepted -->
Department of Surgery, Medical College Kolkata, Kolkata, India.

##### Anurag Mishra (AM) <!-- Accepted -->
Department of Surgery, Maulana Azad Medical College, New Delhi, India.

<!--
##### Priti Patil (PP)
WHO collaboration Centre for Research in Surgical Care Delivery in LMIC, Mumbai, India.
-->

##### Arun Prasad (AP) <!-- Accepted -->

Department of Surgery, Indraprastha Apollo Hospitals, New Delhi

<!--
##### Amit Roy (AR)
Department of Surgery, Sir Nil Ratan Sircar Medical College & Hospital, Kolkata, India.
-->

<!--
##### Siddarth David (SD)
Department of Global Public Health, Karolinska Institutet, Stockholm, Sweden.

Doctors For You, Mumbai, India.
-->

##### Harris Solomon (HS) <!-- Accepted -->
Department of Cultural Anthropology and the Duke Global Health Institute, Duke University, Durham, USA.

##### Megha Tandon (MT) <!-- Accepted -->
Department of Surgery, Vardhman Mahavir Medical College and Safdarjung Hospital, New Delhi, India.

## Keywords
Trauma management, Accident and emergency medicine, Education and training.

## Word Count
2836

## Role of study sponsor and funders
The funding sources had no role in the design of this study nor during its execution, analyses, interpretation of the data, or decision to submit the results.

<!-- \newpage -->

# Abstract

**Objective** To assess the feasibility of conducting a cluster randomised controlled trial comparing the effect of Advanced Trauma Life Support^®^ (ATLS^®^) and Primary Trauma Care (PTC) with standard care on patient outcomes.

**Design** A pilot pragmatic three-armed parallel, cluster randomised, controlled trial between `r start.date` and `r end.date`. Patient follow up was 30 days.

**Setting** Tertiary care hospitals across metropolitan areas in India.

**Participants** Adult trauma patients and residents managing these patients.

**Interventions** ATLS^®^ or PTC training for residents in the intervention arms.

**Main Outcomes and Measures** The outcomes were consent rate, lost to follow up rate, pass rate, missing data rates, differences in distribution between observed and data extracted from medical records as well as all cause and in-hospital mortality at 30 days from the time of arrival to the emergency department.

**Results** Two hospitals were randomised to ATLS^®^, two to PTC, and three to standard care. We included `r n.patients` patients and `r n.residents` residents. The percentage of patients who consented to follow up was `r p.consent`% and the percentage of residents who consented to training was `r recruitment.rate.residents`%. The lost to follow up rate was `r rate.lost.to.follow.up`%. The pass rate was `r pass.rate`%. The missing data was overall low for key variables. Data collected through observations were similar to data extracted from medical records, but there was more missing data in the extracted data. `r number_to_text(m30d.standard.care.n, first = TRUE)` `r m30d.standard.care.p` patients died within 30 days in the standard care arm, `r number_to_text(m30d.atls.n)` `r m30d.atls.p` patient in the ATLS^®^ arm, and `r number_to_text(m30d.ptc.n)` `r m30d.ptc.p` patients in the PTC arm.

**Conclusions** Conducting a full-scale cluster randomised controlled trial comparing the effects of ATLS^®^, PTC, and standard care on patient outcomes will be feasible, especially if such a trial would use data and outcomes available in medical records.

**Trial Registration** ClinicalTrials.gov (reg. no NCT05417243)

# Strengths and limitations of this study

- Prospective data collection with direct observations by dedicated research officers.
- Lack of a priori defined success criteria and thresholds for the feasibility outcomes.
- Use of sealed envelopes potentially compromised allocation concealment.
- Participating centers' heterogeneity may affect the study estimates and bias the results.

\newpage

<!--
# Article Summary

Strengths and limitations of this study:

- Cluster randomized controlled trial comparing the effect of ATLS^®^, PTC and standard care on patient and provider outcomes.
- Prospective data collection with direct observations by dedicated research officers.
- Participating centers' heterogeneity may affect the study estimates and bias the results.
-->

# Introduction

Trauma, defined as the clinical entity composed of physical injury and the body's associated response, causes 4.3 millions deaths every year [@injuries2020]. Several trauma life support training programs have been developed to improve the early management of patients in hospital by providing a structured framework to assessment and treatment [@Mohammad2013; @Jayaraman2014; @Kadhum2020].

The proprietary Advanced Trauma Life Support^®^ (ATLS^®^) and the low-cost alternative Primary Trauma Care (PTC) are two widely established trauma life support training programmes with over a million physicians trained in over 80 countries [@acsAtls2018; @ptcfPtc2018]. Observational studies indicate that these programmes may improve patient outcomes [@vestrup_impact_1988; @Ariyanayagam1992; @ali_trauma_1993; @van_olden_clinical_2004; @VanHeng2008; @Wang2010; @drimousis_advanced_2011; @noordin_evolution_2011; @hashmi_hospital-based_2013; @Petroze2014; @magnone_impact_2015; @bellanova_mountain_2016; @CioèPeña2016], but there is no high quality evidence from controlled trials to support this [@Mohammad2013; @Jayaraman2014; @Kadhum2020; @Jin2021; @putra_impact_2023; @mciver_effect_2024].

Several studies, including at least two randomised studies [@Ali1996; @Ali1999], show that ATLS^®^ is associated with improved knowledge and skills among providers [@Mohammad2013]. Observational evidence suggests that PTC also leads to improved provider skills [@Kadhum2020]. The missing link is then whether these improved knowledge and skills translate into measurably improved patient outcomes.

Systematic reviews call for controlled trials in settings where these programmes are not routinely implemented [@Jayaraman2014; @Kadhum2020], because conducting such effectiveness trials in settings where they are part of the standard of care is not possible. Many settings without routinely implemented trauma life support training are in low- and middle income countries, where trial logistics can be more challenging.

We therefore conducted a pilot study to assess the feasibility of a cluster randomised controlled trial comparing the effect of ATLS^®^ and PTC with standard care on outcomes in adult trauma patients.

<!-- add citation to scoping review of TQIP [@mciver_effect_2024]-->

# Methods

## Protocol Deviations

Protocol deviations are mentioned where relevant in this manuscript, but a list of all deviations is also included as Supplementary Material S1 for completeness.

## Trial Design

We piloted a three-armed cluster randomised controlled trial [@GerdinWärnberg2022]. There were a standard care arm and two intervention arms, ATLS^®^ and PTC training. We collected data for four months in all three arms, first during a one month observation phase and then during a three month intervention phase (or continued observation in the standard care arm). In the published protocol, we also aimed to estimate probable effect sizes and other measures needed for the sample size calculations of a full-scale trial [@GerdinWärnberg2022], but we revised this aim in the light of current guidance on the conduct and reporting of pilot trials [ADD CITATION].

## Study Setting

We conducted this pilot study in seven tertiary hospitals across metropolitan areas in India, where neither ATLS^®^, PTC, nor any other established trauma life support training program is routinely taught. Details about each cluster is included as Supplementary Table S1. Originally, we intented to include six hospitals as clusters, but decided to include a seventh when they expressed interest and as we had the budget to accommodate this. These seven hospitals were a convenience sample of hospitals in India that fulfilled the inclusion criteria, and with existing connections to the research team.

<!--
The trauma system is still developing, with
limited prehospital care, and the in hospital trauma mortality as well
as the proportion of preventable deaths remain high. Lack of standard
trauma training for healthcare providers, limited hospital resources,
inadequate processes of care, overcrowding emergency departments - are
some of the factors that contribute to the high mortality and
morbidity. During recent years efforts have been made to improve
hospital trauma care, through capacity building for trained trauma
care providers, augmenting facilities, and developing care protocols
within the hospitals. 
-->

## Standard Care

Standard care varies across hospitals in India, but most surgical and emergency medicine departments in India organise their physicians in units. These units include both faculty members and residents, who are assigned a specific day of the week when they are posted in the emergency department. In the emergency department, trauma patients are initially assessed by either surgical or emergency medicine residents who also resuscitate patients, perform interventions and refer patients for imaging or other investigations. Compared with other settings where a trauma team approach is adopted, nurses and other healthcare professionals are only involved to a limited extent during the initial management. We did not collect data on how standard care varied between the participating hospitals.

## Intervention

In each intervention arm the residents in one or two units were trained in either ATLS^®^ or PTC. For the purpose of this pilot study, our target was to train a minimum of 75% of residents in each unit. We did not train the units' faculty, because they are typically not directly involved in the initial management of trauma patients. The ATLS^®^ training was conducted in an ATLS^®^ certified training centre in Mumbai and the PTC training was conducted in New Delhi. Both trainings were conducted according to their respective standard curriculum [@acsAtls2018; @ptcfPtc2018], and we did not modify or adapt the delivery or content of these programmes during this pilot study.

The provider courses of both programmes are 2.5 days, and they teach how to assess, resuscitate and stabilize trauma patients and how to adapt to needs of different patient populations, for example pediatric, pregnant and geriatric patients. The teaching is based on case discussions and skill stations. There are some important differences between the two programs. The ATLS course include more on the inter-hospital transfer of patients and includes a greater emphasis on the trauma team [@acsAtls2018]. The PTC course on the other hand focuses on trauma care in the low resource setting [@ptcfPtc2018].

## Eligibility Criteria for Cluster and Participants

### Hospitals

We included tertiary care hospitals in metropolitan areas in India that admitted more than 400 adult patients with trauma each year, and that had operation theatres, X-ray, CT, and ultrasound facilities, and blood bank available around the clock.

### Clusters

We defined a cluster as one or more units of physicians providing trauma care in the emergency department of Indian tertiary care hospitals. To be eligible, units could have no more than 25% of their physicians trained in either ATLS^®^, PTC, or similar training programs before the start of the pilot study. Those residents who had received training in the last five years were considered as trained. The figure of 25% was decided through consensus in the research team, to balance feasibility and contamination of results. The principal investigator at each hospital selected the units for training. We randomised on the hospital level to avoid contamination between intervention arms and the standard care arms.

### Residents

We trained resident doctors doing their speciality training in surgery or emergency medicine managing trauma patients in the emergency department and who were expected to remain in the participating hospitals for at least one year from the time of the training. Consent was sought from the residents in each of the intervention groups before they underwent the ATLS^®^ or PTC training. In the published protocol, we stated that only surgical residents would be trained. However, in some of the participating hospitals, emergency medicine residents performed the initial resuscitation and management of trauma patients, and we therefore decided to train them instead.

### Patients

We included persons who were 15 years or older and presented to the emergency department at participating hospitals with a history of trauma when a designated unit was on duty. History of trauma was defined as having any of the external causes of morbidity and mortality listed in block V01-Y36, chapter 20 of the International Classification of Disease version 10 (ICD-10) codebook as reason for presenting.

## Feasibility outcomes

Our feasibility outcomes were:

- Consent rate of patients and residents. This was equal to the percentage of patients or residents who consented to be included, out of the total number of eligible patients or residents. 
- Lost to follow up rate. This applied only to patients and was equal to the percentage of patients who did not complete 30 day follow up, out of all enrolled patients.
- Pass rate. This applied only to residents in the intervention arms and was equal the percentage of residents who passed the training programme, out of the total number of trained residents.
- Missing data rate. This applied to each outcome and variable and was equal to the percentage of missing values.
- Differences in distributions between directly observed data and data extracted from medical records. This applied to all variables that could be reasonably expected to be present in the medical records, but we only extracted this data from a convenience sample of patients to reduce the workload of the project officers collecting data.

We did not specify success criteria for these outcomes.

<!--The primary outcome and most secondary outcome were assessed and compared both as final values and as change from baseline. All outcomes that pertain to the individual participant level are detailed in Supplemental Material 1. We decided to include a large number of outcomes, including some more exploratory, so that we could test their feasibility and relevance. -->

## Participant Timeline and Inclusion

### Patients

Arriving patients were screened for eligibility and consented, if conscious. Unconscious patients were consented by the patient’s representative. This proxy consent was reaffirmed by the patient, on regaining consciousness. Patients who did not regain consciousness were included based on their representative's consent. We followed up patients at 24 hours after arrival at the emergency department, and at 30 days after arrival at the emergency department.

### Residents

Participating units were screened for eligibility once hospitals confirmed their participation. All residents in these units were approached to consent to training if their hospital was randomised to either of the intervention arms. The protocol stated that residents would be approached for consent before the hospitals were randomised, but this proved not to be feasible, and we therefore asked residents for consent after the hospitals were randomised but before training. The training was conducted approximately one month after the study started in that hospital. Initially, we planned to use simple random sampling to select the units to be trained in each intervention hospital, but for pragmatic reasons we decided to leave the decision on which units to train to the site principal investigator.

## Sample size

We wanted to include at least two clusters per arm to avoid drawing conclusions based on single centers, and we wanted to train at least two units per intervention cluster to be able to evaluate the logistics of sending residents for training. We did not conduct a formal power calculation for this pilot study, as the purpose was to assess the feasibility of the trial logistics and research methods. <!--The time period was dictated by budget and time constraints.-->

## Allocation and blinding

We used simple randomisation implemented using sealed envelopes to allocate sites to trial arms. We did not blind investigators, residents or patients to the intervention.

## Data Collection

Data was collected over a four-month period. A research officer collected data on all patients who presented on the days and shifts when participating residents were assigned to trauma care. The research officers observed care and interviewed residents and patients, and also extracted data from the hospital records. We followed up admitted patients for their complications and other in-hospital outcome measures. Patients who were not admitted or who were discharged before the end of the study were followed up telephonically for mortality outcomes and quality of life outcomes.

## Variables

The research officers collected data on demographics, vital signs, management details including imaging and surgery, and details of any injury sustained. All injuries were coded according to the International Classification of Diseases version 10 (ICD-10). Based on these ICD-10 codes, we calculated the Injury Severity Score using the R package `icdpicr` [@icdpicr_2022]. We also collected data on potential outcomes for the full-scale trial, including mortality within 30-days and in-hospital, complications and cause of death. A list of variables is available in Supplementary Table S2.<!--Supplemental Material 2 contains a full variable list, with definitions.-->

## Patient and public involvement

We conducted community consultations to collect inputs from patients, their caregivers, patient groups, and resident doctors to be used in the selection of outcome measures and implementation of the full-scale trial. The results of these consultations are published separately [@david_patient-reported_2024]. We initially planned to distribute periodic surveys to residents and follow them up after 30 days after training, but we changed this to interviews after the end of the study period as we thought this would provide richer data (not published).

<!--
## Data management
We used the online data collection tool KoboToolbox [@KoboToolbox], accessible only over a virtual private network (VPN), for each participating hospital to upload pseudonymised data to secure servers. Data validation techniques such as restricted values or values within a specific range were used to avoid ambiguous data entries and ensure the validity of the data. Ambiguous responses and errors were resolved after discussion with the core team during weekly meetings. An instruction manual or codebook for data variables was prepared to ensure consistency in data entry. This manual was referred to during the project data collection and variable descriptions were visible for each variable in the online data collection tool. Pseudonymised data were stored at the centralised server. The data were only accessible by the project's principal investigator or by delegation by the project principal investigator.
-->

## Data monitoring

We conducted weekly online meetings to monitor the study and data collection. We conducted one interim analysis approximately halfway through the study, and decided to complete the study as residents and patients were consenting to be included in the study and key variables including mortality outcomes could be collected. We did not use a data monitoring committee.

## Statistical Methods

We used the `r version$version.string` Statistical Software for all analyses [@R_2024]. We analysed all data, including the feasibility outcomes, using descriptive statistics and did not perform any formal hypothesis tests [@Lee2014]. Initially, we planned to analyse feasibility outcomes both on an overall and individual cluster levels, but the sample sizes in individual clusters were too small to generate meaningful results. Quantitative variables are summarised as median and interquartile range. Qualitative variables are presented as absolute numbers and percentages. Additional analyses performed according to the original protocol are described in the Supplementary Material S2 and their results shown in Supplementary Material S3 and Supplementary Tables S1-12.

## Ethics and Dissemination

We were granted research ethics approval from the institutional ethics committees at each participating hospital. For each participating hospital, the approvals were HBTMC/266/SURGERY for Dr R N  Cooper Municipal General Hospital in Mumbai, IEC(II)/OUT/134/2022 for Seth GS Medical College and KEM Hospital in Mumbai, ICC/214/22/20/05/2022 for Lokmanya Tilak Municipal Medical College and General Hospital, CREC/2022/FEB/1(ii) for MEDICA Superspeciality Hospital in Kolkata, MC/KOL/IEC/NON-SPON/1217/11/21 for Medical College, Kolkata, NRSMC/IEC/93/2021 for Nilratan Sircar Medical College & Hospital in Kolkata, and finally IEC-03/2022-2332 for the Postgraduate Institute of Medical Education and Research, Chandigarh.

# Results

We enrolled `r n.patients` trauma patients from `r number_to_text(n.centres)` participating centres between `r start.date` and `r end.date`. The standard care arm enrolled `r n.control` patients, the ATLS^®^ arm enrolled `r n.atls` patients, and the PTC arm enrolled `r n.ptc` patients. We trained a total of  `r number_to_text(n.residents)` residents, `r number_to_text(n.atls.residents)` in ATLS^®^, and  `r number_to_text(n.ptc.residents)` in PTC.

The study flow diagram is shown in Figure \@ref(fig:flow-diagram) and patient sample characteristics across trial arms are shown in Table \@ref(tab:sample-characteristics). Overall, the number of females were `r females`, the median (IQR) age was `r median.age` years, and the median ISS (IQR) was `r median.iss`. A total of `r m30d` patients died within 30 days after arrival to the emergency department, and `r inhosp` patients died in hospital.

After training, a total of `r m30d.standard.care` patients in the standard care arm died within 30 days, compared to `r m30d.atls` patients in the ATLS^®^ arm and `r m30d.ptc` patients in the PTC arm. The corresponding figures for in-hospital mortality were `r inhosp.standard.care`%, `r inhosp.atls`%, and `r inhosp.ptc`% for the standard care, ATLS^®^ and PTC arms respectively.

```{r flow-diagram, echo = FALSE, fig.cap="Study flow diagram. Abbreviations: ATLS, Advanced Trauma Life Support; PTC, Primary Trauma Care.", out.width="30%", out.height="30%"}
knitr::include_graphics("consort-flow-diagram.png")
```

```{r sample-characteristics, echo = FALSE, results = "asis", tab.cap = "Patient sample characteristics"}
sample.characteristics.table |>
  modify_caption("Patient sample characteristics")
```

## Outcomes

The percentage of patients who consented to follow up was `r p.consent`% and the percentage of residents who consented to training was `r recruitment.rate.residents`%. Among patients, the lost to follow up rate was `r rate.lost.to.follow.up`%. Among residents, the pass rate was `r pass.rate`%. The missing data rate ranged from `r min.missing.data` to `r max.missing.data`%, with details for selected variables shown in Table \@ref(tab:sample-characteristics) and in Supplementary Table S2. The variables with the maximum amount of missing data were in the cost of treatment, reported in Supplementary Tables S1-12.

The differences in distributions between observed data and data extracted from medical records, for selected variables that were collected through observation or interview, are shown in Table \@ref(tab:observed-vs-extracted). Overall, the data were similarly distributed, but there were considerably more missing values in the data extracted from medical records compared to the directly observed data.

```{r observed-vs-extracted, echo = FALSE, tab.cap = "Differences in distributions between directly observed data and data extracted from medical records, for selected variables that were collected through observation or interview."}
observed.vs.extracted.comparison.table
```

# Discussion

We show that it is feasible to conduct and collect data for a cluster randomized controlled trial comparing ATLS^®^ with PTC and standard care. Missing data were low for key variables, including the primary outcome and many secondary outcomes. Some variables, especially the cost of treatment (reported in Supplementary Table S2-13) had very high missing data rates and may not be feasible to include in a full-scale trial, or require different data collection methods. The missing data was substantially higher when data was extracted from medical records instead of being directly observed, but the data were similarly distributed, indicating that data collected from medical records is reliable even if it is less complete.

<!-- Should we include dependency at discharge, for example using the
Modified Oxford Handicap Scale used in @Roberts2013?-->

All-cause 30-day mortality was missing in `r rate.lost.to.follow.up`% of patients. This may appear high, especially compared to for example the CRASH-2 and REACT-2 trials, which report missing primary outcome in less than 0.01% of patients [@Roberts2013; @Sierink2016]. Like many other trauma trials, both CRASH-2 and REACT-2 used in-hospital mortality as their primary outcome measure, whereas we attempted to follow up patients after discharge. Our missing data rate for in-hospital mortality was only `r p.missing.in.hospital.mortality`%, which is comparable to previous trials. The difference in the number of patient deaths between the arms may have resulted from the randomisation process with a small number of heterogeneous clusters, and this heterogeneity highlights the importance of taking varying cluster sizes into account in the design of the full scale trial.

<!-- Consider discussing non-inferiority vs superiority. -->

During the course of this pilot study we deviated from the protocol in several ways. We have reported these deviations were appropriate in the methods section, and provide a detailed list as Supplementary
material S1. The most important deviation was the revision of the aim of this pilot study, because we initially planned to estimate potential effect sizes and other measures to help sample size calculations for a full scale trial, in addition to the feasibility outcomes. However, current guidance advice against using pilot studies to estimate effect sizes, as the usefulness of these estimates is questionable [INSERT REF]. We therefore decided to only report patient outcomes descriptively. Another important deviation from the protocol was the training of emergency medicine residents. We initially planned to train only surgical residents, but trauma management routines varied between participating hospitals, and we decided to adapt to local routines. The full scale study will need to accommodate this variation as part of the protocol.

There are some important limitations of this pilot study and therefore lessons to be learned and factored into the design of the full-scale trial. First, the volumes of patients included at some of the participating hospitals were lower than expected. A careful assessment of patient volumes as part of the screening process will be needed for the full-scale trial. Second, data on complications and cause of death were almost universally missing. To collect data on these variables will require other methods, as these were not explicitly described in the medical records and autopsy reports are not readily available. Third, we did not collect in-depth data on the standard care at each hospital. This data should be collected as part of the screening process for the full-scale trial.

<!-- Consider discussing waiver of informed consent for initial 
data collection -->

Previous studies on the effect of ATLS^®^ or PTC training on patient outcomes are observational or quasi-experimental without a control group, with heterogeneous results [@vestrup_impact_1988, @Ariyanayagam1992; @ali_trauma_1993; @van_olden_clinical_2004; @VanHeng2008; @Wang2010; @drimousis_advanced_2011; @noordin_evolution_2011; @hashmi_hospital-based_2013; @Petroze2014; @magnone_impact_2015; @bellanova_mountain_2016; @CioèPeña2016]. Most studies have found that these programmes are associated with improved outcomes, although not all studies have found significant effects [@vestrup_impact_1988; @ali_trauma_1993; @van_olden_clinical_2004; @Wang2010; @noordin_evolution_2011; @hashmi_hospital-based_2013; @Petroze2014; @magnone_impact_2015; @bellanova_mountain_2016]. In contrast, some studies have found that these programmes may be associated with increased mortality [@drimousis_advanced_2011; @CioèPeña2016].

Considering the widespread use of trauma life support training, several systematic reviews call for trials in settings where these programmes are not routinely implemented [@Mohammad2013; @Jayaraman2014; @Kadhum2020]. Our study represents the first published attempt at a controlled trial of the effect of trauma life support training on patient outcomes, and we conclude that conducting a full-scale cluster randomised trial should be feasible after incorporating the lessons of this pilot.

# Contributorship statement

MGW conceived of the study, performed the analysis and drafted and revised the manuscript. AG, AM, CJ, DKV, HS, JB, KDS, LFT, LS, MH, MK, NR, PB, PP, RS, SD, and VK contributed to the design of the
study. MGV, DKV, KDS, and MK drafted the first version of the
protocol. AG, HS, and SD drafted the first version of the patient and
public involvement activities. JB and PP drafted the first versions of
the data management sections and wrote the data management plan. PB
and PP drafted the first versions of the statistical analysis
section. AG, AM, CJ, DKV, HS, JB, KDS, LFT, LS, MH, MGW, MK, NR, PB,
PP, RS, SC, SD, and VK contributed to the refinement of the
protocol. DB, JB, SC, LFT, GG, MK, TK, CJ, NR, RS, KDS, LS and VP interpreted the results and revised the manuscript. AR, AC, C, DK, GG, MK, MT, VK and VP are representatives of participating hospitals.

# Competing Interests

Several authors are ATLS^®^ and/or PTC instructors.

# Funding

Doctors for You through grants awarded to Karolinska Institutet by the Swedish Research Council (grant number 2020-03779) and the Laerdal Foundation (grant number 2021-0048).

# Data Sharing Statement

The code for analysis is released publicly. The final anonymized dataset is available from the corresponding author on request.

\pagebreak

# References