### 2025-07-23 supplementary-outcome-tables.Rmd

\pagebreak

# Results with 75% confidence intervals (CI)

## Additional Online Table 1. Outcomes in all patients during the entire study period, by treatment arm (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Additional Online Table 2. Outcomes in all patients before training, by treatment arm (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 3. Outcomes in all patients after training, by treatment arm (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 4. Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 5. Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 6. Absolute and relative differences in outcomes after training, comparing standard care with ATLS (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 7. Absolute and relative differences in outcomes after training, comparing standard care with PTC (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 8. Absolute and relative differences in outcomes after training, comparing ATLS with PTC (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Additional Online Table 9. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 10. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 11. Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC (75% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.75"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak

\pagebreak

# Results with 85% confidence intervals (CI)

## Additional Online Table 1. Outcomes in all patients during the entire study period, by treatment arm (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Additional Online Table 2. Outcomes in all patients before training, by treatment arm (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 3. Outcomes in all patients after training, by treatment arm (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 4. Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 5. Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 6. Absolute and relative differences in outcomes after training, comparing standard care with ATLS (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 7. Absolute and relative differences in outcomes after training, comparing standard care with PTC (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 8. Absolute and relative differences in outcomes after training, comparing ATLS with PTC (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Additional Online Table 9. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 10. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 11. Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC (85% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.85"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak

\pagebreak

# Results with 95% confidence intervals (CI)

## Additional Online Table 1. Outcomes in all patients during the entire study period, by treatment arm (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients during the entire study period, by treatment arm"]]

```

\pagebreak

## Additional Online Table 2. Outcomes in all patients before training, by treatment arm (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients before training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 3. Outcomes in all patients after training, by treatment arm (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Outcomes in all patients after training, by treatment arm"]]

```

\pagebreak

## Additional Online Table 4. Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 5. Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Relative change from baseline for all outcomes, comparing the period after training with the period before training, by treatment arms"]]

```

\pagebreak

## Additional Online Table 6. Absolute and relative differences in outcomes after training, comparing standard care with ATLS (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 7. Absolute and relative differences in outcomes after training, comparing standard care with PTC (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 8. Absolute and relative differences in outcomes after training, comparing ATLS with PTC (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in outcomes after training, comparing ATLS with PTC"]]

```

\pagebreak

## Additional Online Table 9. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with ATLS"]]

```

\pagebreak

## Additional Online Table 10. Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing standard care with PTC"]]

```

\pagebreak

## Additional Online Table 11. Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC (95% CI)

Empty cells indicate that there were no observations to calculate the statistic.

```{r, echo = FALSE}

outcome.results.tables[["ci.level.0.95"]][["Absolute and relative differences in changes from baseline for all outcomes, comparing ATLS with PTC"]]

```

\pagebreak